Pesticides in small agricultural streams in Germany

### 2016-08-19 article.tex
Update on Overleaf.

\citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
However, this study reflected only a small part of data that is available from national monitoring programs.
These programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
Large amounts of data are generated, which possibly can also be used to answer other questions.
In Germany monitoring programs are setup independently by the federal states in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
However, currently there is no curated national-wide compilation of this data available.

@@

All data-processing and analyses have been performed using R \citep{r_core_team_r:_2016}.
To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.

@@

where $No_i$ is the observed number of exceedances at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site. 
We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) and used the number of sampling events per site ($n_i$) as an offset to account different sampling efforts. 
$f_1$ and $f_2$ are smoothing functions using thin plate regression splines \citep{wood_thin_2003}.
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement for details). 
We used pointwise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

@@

The compiled dataset comprised only few standing waters (58 sites) and the majority (90\%) of samples where taken via grab sampling.  % see clean.R for numbers
Therefore, we report only results of grab samples from streams. 
The analysed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).

@@

In total 484 different compounds used as pesticides and their metobalites were measured at least once (Supplement, Table \todo{Set ref to table}). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
We found substantial differences of the spectra of analysed compounds between federal states (Figure \ref{fig:fig2}).
Hierarchical clustering revealed three groups:
i) with less then 100 compounds (SL, ST and TH), \\
ii) with a medium sized spectra and \\
iii) with a big and distinct spectra (RP and NI).

@@

\subsection{Pollution of small agricultural streams}
Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
%EQS
Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:fig5}A).
% RAC
Neonicotinoid Insecticides and Chlropyrifos showed highest risk quotients.
For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
In 15\% of all samples (n = 6377) risc quotients higher then 1 were observed.
Highest RQ were observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
% TU
In 33\% of the samples no pesticides were detected. 
The mean $log(TU_{max})$ for detects was -4.6.
2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
We found a high correlation of $log(TU_{max})$ and $TU_{sum}$ \todo{ausrechnen und r angeben}, with a maximum higher value of $TU_{sum}$ xx units.
%Mixtures
In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:fig5}D).

@@

% current problems in monitoring and possible solutions
A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analysed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.

@@

% TU
xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshhold above biological effects have been shown \todo{add citation schaefer}. 
The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with one dominant compound.
% Mixtures
Nevertheless, most pesticides did not occur individually but in mixtures.
This highlights that

>>>>

\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
However, this study reflected only a small part of data that is available from national monitoring programs.
These programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
Large amounts of data are generated, which possibly can also be used to answer other questions.
In Germany monitoring programs are setup independently by the federal states in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
However, currently there is no curated national-wide compilation of this data available.

@@

All data-processing and analyses have been performed using R \citep{r_core_team_r:_2016}.
To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.

@@

where $No_i$ is the observed number of exceedances at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site. 
We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) and used the number of sampling events per site ($n_i$) as an offset to account different sampling efforts. 
$f_1$ and $f_2$ are smoothing functions using thin plate regression splines \citep{wood_thin_2003}.
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement for details). 
We used point-wise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

@@

The compiled dataset comprised only few standing waters (58 sites) and the majority (90\%) of samples where taken via grab sampling.  % see clean.R for numbers
Therefore, we report only results of grab samples from streams. 
The analyzed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).

@@

In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table \todo{Set ref to table}). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
We found substantial differences of the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
Hierarchical clustering revealed three groups:
i) with less then 100 compounds (SL, ST and TH), \\
ii) with a medium sized spectra and \\
iii) with a big and distinct spectra (RP and NI).

@@

\subsection{Pollution of small agricultural streams}
Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
%EQS
Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:fig5}A).
% RAC
Neonicotinoid Insecticides and Chlropyrifos showed highest risk quotients.
For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
In 15\% of all samples (n = 6377) risk quotients higher then 1 were observed.
Highest RQ were observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
% TU
In 33\% of the samples no pesticides were detected. 
The mean $log(TU_{max})$ for detects was -4.6.
2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
We found a high correlation of $log(TU_{max})$ and $TU_{sum}$ \todo{ausrechnen und r angeben}, with a maximum higher value of $TU_{sum}$ xx units.
%Mixtures
In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:fig5}D).

@@

% current problems in monitoring and possible solutions
A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.

@@

% TU
xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshold above biological effects have been shown \todo{add citation schaefer}. 
The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with one dominant compound.
% Mixtures
Nevertheless, most pesticides did not occur individually but in mixtures.
This highlights that
